
#Assignment Instructions

1.Code for reading in the dataset and/or processing the data

2.Histogram of the total number of steps taken each day

3.Mean and median number of steps taken each day

4.Time series plot of the average number of steps taken

5.The 5-minute interval that, on average, contains the maximum number of steps

6.Code to describe and show a strategy for imputing missing data

7.Histogram of the total number of steps taken each day after missing values are imputed

8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report


knitr::opts_chunk$set(warning=FALSE)

##Loading and Data preparation
 

```{r, echo = TRUE}
library(data.table)
library(ggplot2)

activity <- read.csv("C:/Users/snamin/Documents/GitHub/RepData_PeerAssessment1/activity.csv")

activity$date <- as.POSIXct(activity$date, format ="%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)

summary(activity)

```
What is mean total number of steps taken per day?
 
```{r, echo = TRUE}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "darkblue", ylim = c(0,20), breaks = seq(0,25000, by=2500))

```
What is the mean of the total number of steps taken per day?

```{r, echo = TRUE}

mean(activity_total_steps$steps)

```

What is the median of the total number of steps taken per day?

```{r, echo = TRUE}

median(activity_total_steps$steps)

```
What is the average daily activity pattern?

```{r, echo = TRUE}

average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="darkblue", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")

```
Imputing missing values

```{r, echo = TRUE}

sum(is.na(activity$steps))
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]

# Create a new dataset with the missing data filled in
activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")

# Create its histogram

hist(total_steps_imputed$daily_steps, col = "darkblue", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))

#Calculate mean and median again 

mean(total_steps_imputed$daily_steps)
median(total_steps_imputed$daily_steps)

```
Are there differences in activity patterns between weekdays and weekends?
```{r, echo = TRUE}

#Create a new factor variable in the dataset with two levels - "weekday" and "weekend"

activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Saturday" | weekdays(x) =="Sunday") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```

```{r, echo = TRUE}
#Make plots

activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)




```



